### 2015-08-14 paper/paper.tex
Added Chris' comments v2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.1 Generated 2015/22/05 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass{frontiersHLTH} % for Health articles
%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square}
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}
\linenumbers

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Stewart {et~al.}} %use et al only if is more than 1 author
\def\Authors{Terrence C. Stewart\,$^{1,*}$, Travis DeWolf\,$^{1}$, Ashley Kleinhans\,$^{2}$ and Chris Eliasmith\,$^1$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Centre for Theoretical Neuroscience, University of Waterloo,
    Waterloo, ON, Canada
    \\
$^{2}$Mobile Intelligent Autonomous Systems group, 
      Council for Scientific and Industrial Research,
      Pretoria, South Africa}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Terrence C. Stewart}
\def\corrAddress{Centre for Theoretical Neuroscience, University of Waterloo,
    Waterloo, ON, Canada}
\def\corrEmail{tcstewar@uwaterloo.ca}

\begin{document}
\onecolumn
\firstpage{1}

\title[Embodied Neuromorphic Benchmarks]{Embodied Neuromorphic Benchmarks}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% The sections below are for reference only.
%%%
%%% For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself. It is recommended to organize your manuscript in the
%%% following sections or their equivalents for your field:
%%% Abstract, Introduction, Material and Methods, Results, and Discussion.
%%% Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
%%%
%%%For information about Clinical Trial Registration, please go to http://www.frontiersin.org/about/AuthorGuidelines#ClinicalTrialRegistration
%%%
%%% For Clinical Case Studies the following sections are mandatory: Abstract, Introduction, Background, Discussion, and Concluding Remarks.
%%%
%%% For all other article types there are no mandatory sections.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

    Evaluating the effectiveness and performance of neuromorphic hardware is
    difficult.  It is even more difficult when the task of interest is an
    embodied task; that is, a task where the output from the neuromorphic
    hardware affects some environment, which then in turn affects the
    hardware's future input.  However, embodied situations
    are one of the primary potential uses of neuromorphic hardware.  To address
    this, we present a methodology for generating embodied benchmarks that makes use
    of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
    Minimal simulation has been shown to lead to robust real-world performance,
    while still maintaining the practical advantages of simulation, such as
    making it easy for the same benchmark to be used by many researchers.
    This method is flexible enough to allow researchers to explicitly
    modify the benchmarks to identify specific task domains where particular
    hardware excels.  To demonstrate the method, we present a set of novel benchmarks
    that focus on motor control for an arbitrary system with
    unknown external forces.

\tiny
 \keyFont{ \section{Keywords:} neuromorphic hardware, benchmarking, 
        minimal simulation, adaptive control, neural networks} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\section{Introduction}

Neuromorphic hardware holds great promise for a wide variety of applications.
The combination of massively parallel computation and low power consumption
means that there is the potential to have complex algorithms running in
embedded processing situations, without being a significant drain on available
energy.  A crucial challenge is to identify what sort of always-on or
interactive functionality can best exploit these devices.

To evaluate applications of neuromorphic hardware, we need benchmark tasks.  These
tasks must allow us to compare across different instances of neuromorphic
hardware (and potentially across different algorithms implemented in that
hardware).  Good benchmarks will allow us to quantitatively compare systems, letting
researchers both measure the progress in the field, and also directly
compare competing approaches.

In this paper, we focus on the development of \emph{embodied} benchmarks.
These are dynamic tasks where the output of the neuromorphic hardware \emph{influences
its own future input} through some environment.  This is in contrast to standard categorization or pattern
identification tasks, where the input is some fixed sequence and the hardware
must produce the correct output for each input (or input pattern).

We believe embodied benchmarks should be of particular interest to
neuromorphic research, given that the most compelling applications of neuromorphic
hardware are likely to be in this domain of embedded and interactive control
of robotic or other physical systems.  However, embodiment itself raises a number
of issues that complicate the development of such benchmarks.  Rather than
simply providing a data file of inputs and desired outputs, the benchmark
must either specify a full physical system for that embodiment, or it must 
provide software for a simulation of that system.  As we discuss below, both
approaches are problematic.  
Describing a method for overcoming these shortcomings is the primary goal of
this paper.

\section{Embodied Benchmarks}

An embodied benchmark task is one where the system we are studying has a 
two-way interaction with some sort of environment.  That is, the outputs from
the neuromorphic hardware are sent to the environment where they cause
an effect, the results of which change the subsequent input.  For example,
the outputs might control the movement of a robot, which in turn affects the
sensory data received by the robot.

\subsection{Simulation versus Physical Instantiation}

To define an embodied benchmark, we need to be explicit about the embodiment.  If
a robot is to be controlled, we need to specify all of the details of that robot.  What motors does it have?  How are they
configured?  How strong are they?  What sensors are there?  Where are they
placed?  How accurate are they?  However, even if these questions are
answered, there is a fundamental problem in that \emph{other researchers
need access to that exact robot}.  If a benchmark is to be widely used, other
researchers developing their own neuromorphic hardware should be able to do their own 
testing on the same benchmark system.

Furthermore, using a physical robot imposes significant practical
difficulties when performing extensive benchmark testing.  When testing, we
often want to run the same task over and over again, both for robustness and
to see the effects of varying parameters.  With a physical robot, this means
manually setting up the task, letting the test run, gathering the resulting
data, and then resetting the robot back to the initial state.  Consequently,
issues like battery life become problematic, and not just because there is a
limited amount of time available for testing.  As the battery level changes,
the performance of the robot itself can also change.  Futhermore, for any
rigorous testing of the benchmark, we will want to examine situations where
the system fails.  This means that some of the testing will involve parameter
settings that lead to poor behaviour, which might have the undesireable 
result of causing physical damage.

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should we have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing [TODO: ref??].  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap, giving high scores to hardware that does not turn out to
functional well when deployed in real situations.

Furthermore, neuromorphic hardware has another constraint that severely limits
the utility of simulation.  Typically, when a simulation is
too simple to reflect reality, we add details to the simulation itself.
Incredibly finely detailed simulations can be created, filling in all of the
details needed.  However, accurate modelling of physical systems can very quickly
become \emph{impractical to run in real time}.
This is a fundamental problem, in that neuromorphic
hardware is often tied to real-time interactions, and there is no way to slow down
the hardware to match the simulated environment.  This means that even if
we spent the considerable amount of research effort needed to define a
simulated environment for an embodied benchmark, running that simulation 
fast enough to interact with the desired hardware would require more
resources than are available.

[TODO: maybe add a paragraph summarizing the horns of the dilemma so as to set up the
role of minimal simulation more?]

\subsection{Minimal Simulation}

The above considerations seem to indicate that even though using real-world
physical hardware for benchmarking is problematic, it is still better than
using simplistic simulations which may not generalize to real tasks.  However,
we do not think this is the case.  Instead, we believe neuromorphic benchmarking can effectively exploit an approach 
known as \emph{Minimal Simulation}
(\citealt{Jakobi97evolutionaryrobotics}).

This approach was first suggested in the context of evolutionary robotics.  
Notably, the problem faced by embodied neuromorphic benchmarking is remarkably 
similar to that faced earlier by these researchers. 
In evolutionary robotics, the
goal is to use genetic algorithms to \emph{evolve} systems that can control
robots to perform various tasks.  These tasks can be as simple as navigation
and obstacle avoidance, but have also included more difficult tasks such as walking,
object identification, and visual tracking [refs???].

However, performing evolution on real physical robots is problematic
for the same reasons that benchmarks on physical robots are problematic.  The
robots must be reset to the same state each time; they often involve
behaviour that can physically damage the robots; and they take a very long
time to run.  For this reason, attempts were made to evolve algorithms using simulated
robots.  However, the general finding was that algorithms that worked on the
simulated robots would not work when run on the real physical robots.  If
the simulations were improved, adding complex physical detail, then it was
possible to generalize to real behaviour; unfortunately, such complex
simulations would run slower than real-time [refs].

To address this problem, \cite{Jakobi97evolutionaryrobotics} proposed the creation of
``minimal'' simulations.  These are simulations where there is variability
\emph{within the simulation itself}.  In other words, we make \emph{poor}
simulations, but ensure that the way in which they are poor is itself
variable.  We are then in a position to ensure that the controllers work across that whole
range of variability.  ``Instead of trying to eliminate the differences between
simulation and reality, they are acknowledged, and mechanisms are put in place
to prevent evolving controllers from relying on them.'' [ref: jacobi thesis].

With this approach, it became possible to build minimal simulations that would
run faster than real-time and yet also be complex enough that if a system
could successfully control the simulation, it was also likely to successfully
control a real robot.  To achieve this kind of transfer, the simulations were made to be
unreliable in almost every respect.  For example, for a simulation of a simple
motor it would still be the case that if power is applied it would generally
try to spin, but the exact amount of torque, the amount of sensory noise,
the amount of time needed, the amount of static and dynamic friction, and so
on would all be randomly chosen.  A successful controller would have to deal
with this wide range of variability, and if it could handle that variability
then there would be reason to believe it could also handle the real physical system.

It is worth noting that a minimal simulation only has to be a
good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
corridor following behaviour, the dynamics of the simulation might
differ wildly from those of reality if the controller hits a wall or
goes round in circles, but this does
not matter, since the controllers we are interested in transferring
across the reality gap will neither
hit walls nor go round in circles.''  [ref: jacobi thesis]  If the controller is poor, we do not 
need the simulation to be at all accurate in exactly \emph{how} that poor
behaviour is manifest.  We do not need an exact detailed physics model of
the collision between a robot and a wall, or a detailed model of what happens
to a robot arm when it starts oscillating wildly due to a poor control signal.
All we need is for the simulation to be just good enough to indicate that
things have gone wrong, and thus give a low score to that controller.  This
means that, for example, in a minimal simulation of an eight-legged walking robot,
it is not necessary to have a physics simulation that correctly models what
happens when two legs collide with each other.  Rather, if legs collide
with each other, that is an indication that the walking behaviour is very poor.
As long as that result is indicated we can greatly simplify the simulation
by not including all the details necessary to model these complex physical
interactions.

\subsection{Minimal Simulation for Benchmarking}

Given the success of this approach for evolutionary robotics, we propose
using minimal simulation for neuromorphic benchmarks.  We would argue that
one important use of a benchmark is \emph{generalization}.  That is, by knowing how well particular
hardware performs on a benchmark, you can reasonably infer how well
that hardware will perform in other situations.  For example, if an image recognition
algorithm performs well on the MNIST hand-written digit recognition
benchmark, this suggests that it may also perform well on
a different recognition task.  Of course, this inference will fail if that
algorithn has been specifically over-fit to that situtation.  For
that reason, it is useful to have benchmarks that covers a wide range
of variations on the task.  If the hardware performs well across that
variability, then it is more likely to also work in sufficiently similar new situations.

To achieve this kind of transfer, we need software simulations of the environment for the task.
These simulations
must be fast enough to run in real time (so that they can be controlled by
real neuromorphic hardware), and they must be extremely variable, to encourage
robustness of the methods being benchmarked.  Each time
the simulation is run, different parameters will be chosen to give significant variability
(so one run might have a large degree of sensor noise while the next run has
none at all; one run might have more delay in the motor response and another
might have less power available).  Being successful at the benchmark means
being successful across all this variability.

The result is a benchmark that can be used by any researcher.  The fact
that it is a simulation means that source code can be shared, and that no
specialized hardware is needed.  Furthermore, the variability in the simulation
itself can be controlled, and this can help give a rich characterization of
the benchmarked hardware.  For example, some hardware might only work with
small amounts of sensor noise, while other hardware might be most effective when there
is significant delay in the motor response.  This flexibility in parameters
in the benchmark allows researchers to explicitly characterize those particular
situations where their hardware excels.

\subsection{Cheap Robotics}

The minimal simulation described above forms the core of our method for generating benchmarks.
The purpose of these benchmarks is that is that they should do a reasonable job
of generalizing to real-world physical tasks.  Consequently, it is important to
supplement simulation benchmarks with at least one easy-to-construct 
physical analog.  This physical version would be one particular instance of the
type of situation the benchmarks are meant to cover.  For that reason, the physical task
is much more restrictive in terms of what general conclusions can be drawn
from how well different hardware performs in that situation.  Rather, the purpose is to give
an explicit double-check that hardware that performs
well on the simulation benchmarks also perform well in a physical environment.

To keep the physical aspect simple, we recommend cheap, widely-available
components.  This allows a greater chance for other researchers to have
access to the same (or similar) hardware.  For the particular example 
benchmark described in the next section, we use the Lego Mindstorms EV3 kit,
a simple robotics platform available at most toy stores.

It is important to note that there is a theoretical advantage to
using cheap robotics hardware for benchmarking, in addition to the practical
advantages.  In particular, we \emph{do not want benchmarks that rely on
high-speed, high-accuracy devices}.  The purpose of benchmarks
is not to indicate how well this neuromorphic hardware works to control this
particular robot in this task.  Rather, the purpose of a benchmark is to
characterise how well some specific neuromorphic hardware works on a task 
\emph{in general}.  The variability introduced in the minimal simulation means that the hardware
should be able to function across a wide variety of physical embodiments, and
so if we are to choose one particular physical embodiment to test in the
real world, then we should choose one that is not high-precision.
For this reason, we believe using cheap Lego robots is actually more informative
for benchmarking than expensive high-precision robots.\footnote{Of course,
for more complex benchmark tasks we may need sensory and motor capabiities
that are beyond that of a simple Lego robot.}

\section{Example: Adaptive Motor Control}

To demonstrate this approach to creating embodied neuromorphic benchmarks,
we now consider a basic control task.  Suppose we have a system with a number
of joints with positions $q = [q_1, q_2, ..., q_n]$ and we want to send a control signal $u = [u_1, u_2, ..., u_n]$ to the motors at each joint
such that
the joints move to a particular desired position $q_d = [q_{d,1}, q_{d,2}, ..., q_{d,n}]$.  The only output from the controller is
the signal $u$ and the inputs are the current position
of each motor $q$ and the desired positions $q_d$.

The simplest controller for such a situation is a P (proportional) controller,
where $u=K_p(q_d - q)$.  This is often supplemented with a D (derivative) term,
which helps to slow the system down as it approaches the desired position,
thus avoiding overshoot and oscillation. This combination of terms leads to the
standard PD controller $u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$.
Both $K_p$ and $K_d$ are constants that can be
tuned to particular situations.

However, this controller has difficulty in the presence of significant external
forces.  For example, consider a single motor controlling the angle of a single
arm.  If the arm is held out to the side, gravity acting on the mass of the
arm itself will pull the arm downward.  Thus to hold the arm still at a
particular $q_d$ will require the controller to apply a force to counteract
gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
it cannot compensate for gravity, and so the arm will stay stationary at some
angle below the desired angle. [TODO: add diagram]

The standard solution to this steady-state error is to add an I (integral) term ($K_i \int{(q_d-q) dt}$)
to the controller, making it a PID controller.  As the difference between where it is ($q$) and
where we want it to be ($q_d$) accumulates over time, the $K_i$ term will gradually increase the
the extra controlled force $u$ that is being applied until it is large enough to counteract
the external force of gravity (or whatever other external forces are present).
However, this approach has great difficulty when $q_d$ changes, since the
external force due to gravity changes depending on the position of the arm $q$.
The controller ends up having to ``relearn'' the
correct amount of extra force needed every time $q_d$ changes.

In some robotics applications, this problem is solved by mathematically
analyzing the geometry and mass of the system to compute exactly how much extra
force is needed.  In this particular case, the answer is straight-forward,
in that the extra torque due to gravity is $\tau=mg {l \over 2} sin(q)$, where
$m$ is the mass of the arm, $l$ is the length, and $g$ is $9.8m/s^2$.  If the
force applied by the motor is linear in $u$, then we could simply compute this
value and add it to our controller's output.  However, this assumes a perfectly
even distribution of weight in the arm, ignores momentum, friction, and other forces, and 
gets much more complex as more joints are added.  Furthermore, if this
initial computation is slightly off, or if details of the system change,
there is no way to adjust this compensation.

Fortunately, there is an adaptive solution to this problem, and it is one that
fits well with neuromorphic hardware.  \cite{Slotine1987} show that if you
express the influence of the external forces as $\tau=Y(q) \omega$
(where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and $\omega$
is a vector of scalar weights, one for each function in $Y$), then you can learn
to compensate for these external forces by using the learning rule $\Delta \omega = \alpha Y(q) u$,
where $u$ is the basic PD control signal.

Importantly, as pointed out by \cite{SannerSlotine1992} and \cite{Lewis1996}, rather than making explicit
assumptions about the exact functions that should be in $Y(q)$, we can use
a neural network approach where each neuron is a different function of $q$.  
As long as there is enough hetereogenetity (i.e. as long as the neural activity
forms a basis space that is capable of approximating the external forces), then the
learning rule will continue to work.  This approach has been extended to
biologically plausible neurons and been used in both the
Recurrent Error-driven Adaptive Control Hierarchy (REACH) model of human motor control
(\citealt{dewolf2014}) and quadcopter control (\citealt{KomerThesis}).

These considerations suggest that there is a neuromorphic-friendly family of 
algorithms to address the general problem of controlling a wide variety of 
physical systems.  Identifying those algorithms will allow us to benchmark 
their performance across example tasks and physical configurations.  
To implement these algorithms, the input to the
neuromorphic hardware is $q$, the system state.  This input is fed to each
neuron such that each neuron produces some output activity that is based on
this input.  Since $q$ will be multi-dimensional (if there is more than one
joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q + b_i$, 
where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen 
vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities,
or could be learned using some back-propogation of error method.  Here, for simplicity, we only
consider the approach of randomly selecting $e_i$ and $b_i$.},
and $b_i$ is a randomly chosen bias term).  
Given this input, the neurons will produce some output $A$.
We now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
that is initially all zeros.

To use this controller, we add its output to that of the standard PD controller.
That is, the standard controller has $u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$,
and so our actual output to the motor is $u + Ad$.  We then apply a learning
rule on $d$ such that $\Delta d = \alpha A \times u$.  Here, $\alpha$ is a learning
rate and the cross product is used so that we are applying the learning rule
on all the joints simultaneously.

Notice that we can think of this system as a three-layer neural
network, where the input and output layers are linear.  The first layer is $q$, the input state, one value for each joint.  The ``hidden'' layer
is the neurons producing activities $A$, the activity of a large number of neurons.  The output layer again has
one value per joint, and is the extra added signal to apply to the motors, $Ad$.
Given that this is a canonical example of the use of neural networks, we
expect that the majority of neuromorphic hardware is flexible enough to implement
this model.  Importantly, it functions well with spiking neuron models as 
well as non-spiking ones.  For spiking neurons, we consider $A$ to be the
instantaneous measure of the output of a neuron (i.e. whether or not it is currently outputing a spike),
filtered through a low-pass filter.  Further discussion of this sort of learning
rule and comparison to biological spiking neurons is given by [ref?].

\subsection{Online and offline learning}

The rule for modifying the weights $d$ described here is
of a very common form, as the weight
update from a neuron is proportional to the activity of that neuron and an
external error signal.  This makes it an instance of the ubiquitous delta rule.
Thus, neuromorphic hardware that has built-in learning will often be able to natively support this rule.
However, some neuromorphic hardware does not intrinsically have
the ability to update connection weights in this manner.

In that case, there are at least two possible ways to implement this algorithm.
First, the multiplication by $d$ can be done on the
output from the neuromorphic hardware.  Any embodied neuromorphic system will have some method that takes
the neural output from the hardware and sends it to the motors (or to the simulation
of the motors).  Instead of sending the result of $Ad$, the hardware could
send $A$ (the activity of all the neurons), and the interface to the motor
can be responsible for doing the multiplication by $d$ and updating $d$
according to the learning rule.

Alternatively, it may be possible to use offline learning.  That is, rather than updating
the weights $d$ during the simulation, we record $A$ and $u$, and after
a period of time stop the controller, compute the sum of the changes
to $d$, load the new value of $d$ onto the neuromorphic hardware, and
start the controller again.

Given this variety of options for implementing adaptive algorithms of this type, 
we believe it should be possible to benchmark most neuromorphic systems on 
adaptive control tasks in this manner.

\subsection{Minimal Simulation for Adaptive Control}

Now that we have defined the task domain, we can use the principles of minimal
simulation to construct a flexible and variable simulated environment for
testing adaptive control.  In this case, we would like to develop a bare-bones simulation of
the system being controlled, with significant variability.  If the neuromorphic
controller works well across this variability, then it is likely to work well
outside of simulation as well.

The basic system variable is a vector of joint angles $q$.  
Each joint has a velocity $v$.  The force applied by each motor is 
related to the signal $u$ sent to the motors, but will generally have some
maximum value $T$, so we use $tanh(u)T$ to determine the force as a function of the control signal. To account for friction, we 
scale the velocity by some factor $F$ every time step.  
This results in the simplistic simulation described by:

\begin{align}
    \Delta v & =  -v F + tanh(u)T \\
    \Delta q & = v
\end{align}

In addition, we add a external perturbing forces.  In a real system,
this could be the effects of gravity given the current configuration of the
motors, or of other unexpected influences.  
Rather than chosing one particular fixed external force for our benchmark,
we \emph{randomly generate} this force each time the benchmark is run.  This
ensures that the benchmark covers a wide range of possible external forces
and motor configurations, rather than just one particular situation.

Specifically, to generate this force, we start with a small set of
smooth functions $f$ which are often found in dynamics equations ($x$, $x^2$,
$sin(x)$).  We then generate an external force of $K_f (\zeta \cdot f(\beta \cdot q + \gamma) + \eta)$
where $\zeta$, $\beta$, $\gamma$, and $\eta$ are all random vectors and
$K_f$ is a scaling factor to control how strong the external force is.  The
result is added to Equation 1.  For example, if $q$ is 4-dimensional
(i.e. if there are four joints being controlled) and if there are three smooth
functions in $f$, then $\beta$, $\gamma$, and $\eta$ are 
all vectors of length 4 and $\zeta$ is a 4x12 matrix.  To introduce
significant variability, all of these values are
randomly chosen from the normal distribution $N(0,1)$.

To complete the simulation, we introduce additional sources of variability: random noise, delay, and filtering to both the input and the
output of the system.  For noise, we add $N(0,\sigma_u)$ to the control signal
$u$ and $N(0,\sigma_q)$ to the $q$ value reported back to the controller.  We
also use a low-pass filter to smooth both values (with time constants $\tau_u$
and $\tau_q$) after this noise is added, giving a damping effect.  Finally, both $q$ and $u$ are delayed
by an amount of time $t_q$ and $t_u$ to reflect communication delays that are
common in physical systems.

The resulting simulation is not meant to be an accurate portrayal of a particular physical embodiment.
Rather, this simulation is meant to be extremely fast
to simulate, and it is meant to be similarly difficult to control as a real system.
In other words, if a controller manages to be able to control
the various randomly created minimal simulations of embodiment that are
generated with this approach, then we have reason to believe that it will also
be successful at controlling real embodiments.  With this minimal simulation and
modern computers, we can run real-time simulations of systems with dozens of joints that have
highly complex interactions between them.  Consequently, we can effectively
benchmark how well an adaptive
controller deals with these situations.

\subsection{Calibrating the Minimal Simulation via Cheap Robotics}

It is important to ensure that the minimal simulation defined
in the previous section is representative of the sorts of real-world
situations in which we want to use these same controllers.  Importantly, 
this physical instantiation does not have to exactly match any particular
parameter setting of the minimal simulation.  Rather, we want a physical system
that shares basic functional similarities to the minimal simulation defined
previously.

For example, we want the inputs to the system to act like $u$, in that a
positive number will increase some velocity $v$ which will in turn increase some
sensor value $q$.  We want there to be some sort of external applied force that
affects $q$, and we want that external force itself to be a function of $q$.
We want there to be communication delays and noise in the sensor and motor
systems, and we want all of these effects to be somewhere within the
ranges covered by the minimal simulation.  While implementing this kind of
hardware analog cannot guarantee that neuromorphic hardware
that is successful in simulation will be successful in every similar
real-world task, it does provide an existence proof that there is at least
one real-world task where the hardware performs similarly to how it performs in simulation.

For our specific demonstration, we describe an easy-to-build system that can be usefully controlled
by this adaptive method.  In particular, we use the Lego Mindstorms EV3 robot kit,
organized as shown in [TODO: add figure].  It consists of a single
motor, mounted in such a way that there is a significant amount of weight on
the arm itself (i.e., the weight of the motor itself).  Multiple motors can be
added, and other configurations can be considered and should also be suitable
for testing control, but here we consider only this basic case.

To interface to the physical hardware, we installed the \texttt{ev3dev} operating
system (\texttt{http://ev3dev.org}), a Debian-based Linux system specifically developed
for the EV3.  We then installed and ran the \texttt{ev3\_link} program from \texttt{ev3dev-c}
(\texttt{https://github.com/in4lio/ev3dev-c}).  This allows the EV3 to listen for
UDP commands that tell it to set motor values and read sensor values.
Communication with a PC was over a USB link (although the system also
supports WiFi communication).  With constant communication, the system is
able to adjust the power sent to the motors $u$ and give position feedback
$q$ from those motors at a rate of around 200Hz.

Figure \ref{fig:ev3} shows the effects of adaptive control on this physical
system.  Without adaptation (i.e. with a simple PD controller), there system
state $q$ (the joint angle) overshoots the desired $q_d$.  This overshoot
is largest when $q$ is large.  This is because the external force applied to
the joint due to gravity is proportional to $sin(q)$.  The $q$ value also
overshoots and comes back part-way, due to physical momentum.

However, with adaptation (the right-hand side of Figure \ref{fig:ev3}),
the system learns to counteract this extra force due to gravity.  After
the first 5 seconds, the system is able to bring $q$ much closer to the
desired $q_d$.

Now that we have this physical example of the task out minimal simulation
benchmark is meant to cover, we can use it to calibrate the parameters of
the simulation.  For example, communication with the EV3 happens around
200Hz, meaning that there must be a delay on the order of 0.005 seconds.  Given
this, we set the delays in the simulation to be uniformly chosen between
0 and 0.01.  Importantly, we do not need to exactly measure the delay in the
EV3 robot --- we just make sure that the minimal simulation covers even worse
cases.

For sensor noise, we note that the EV3 rotation encoders for the motors (the
devices that measure $q$) have a resolution of 0.0175 (1 degree).  This is a
very different sort of noise than the gaussian noise used in the simulation,
so we set the simulation noise $\sigma_q$ to be much larger (uniformly distributed
between 0 and 0.1).  Similarly, the motor resolution is 0.01, as it accepts 
integer values up to 100, so we set the motor noise $\sigma_u$ to be uniform between
0 and 0.1.

Finally, we can use the physical system to calibrate the relationship between
$T$ (the maximum torque applied by the motor) and $K_f$ (the scaling factor
of the external force).  After all, we do not want external forces that are
so strong that the system does not have enough strength to counteract them.
On the physical robot, in the worst-case scenario ($q=\pi/2$ or $-\pi/2$),
the motors must be driven at around 0.3 times their maximum strength to
balance the force of gravity.  (This can, of course, be adjusted by changing
the weight and its position on the end of the arm).  If we arbitrarily fix $K_f$ to 1 and randomly
generate external forces given the process described above, then 95\% of
the time we get values between -3.75 and +3.75.  Since we want the motors to
be strong enough to compensate for forces in that range, we set $T$ to 10.

\section{Benchmark Analysis}

- what do we measure?   performance: rmse and delay.  Also, power consumption?  cost?

- for given hardware, how many neurons can be done in realtime?  that's how many should be used for comparison

- show how these things change as we change parameters in the simulation

- delays

- number of motors $N$

- magnitude of nonlinearity $K_f$

CPU, GPU, and SpiNNaker?

\section{Other Benchmarks}

(sketch out how this approach could be applied to other task)

Adaptive Jacobian

Driving down a corridor, avoiding obstacles

Classical conditioning

Operant conditioning

\section*{Disclosure/Conflict-of-Interest Statement}

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Acknowledgments}
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

\textit{Funding\textcolon} Text Text Text Text Text Text  Text Text.

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
\bibliography{paper}

%%% Upload the *bib file along with the *tex file and PDF on submission if the bibliography is not in the main *tex file

\section*{Figures}

%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.jpg and logo2.eps) are in the Frontiers LaTeX folder
%%% If using *.tif files convert them to .jpg or .png

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/fig_ev3}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:ev3} Figure \arabic{figure}.}{ Adaptive control of the EV3 lego robot used for calibrating the minimal simulation.
 The effects of adaptation over two different desired trajectories are shown.  Without adaptation, the joints $q$ do not reach the desired $q_d$ when $q_d$ is large (which
 is when the external force is largest).  With adaptation, $q$ is closer to $q_d$ after about 5 seconds, showing that the system has quickly learned to compensate.}
\end{figure}

%\begin{figure}
%\begin{center}
%\includegraphics[width=10cm]{logo2}% This is an *.eps file
%\end{center}
%\textbf{\refstepcounter{figure}\label{fig:02} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
%\end{figure}

%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.

%%% Frontiers will add the figures at the end of the provisional pdf automatically %%%

%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.

\end{document}